---
title: "quartodoc"
format: html
editor: visual
warning: FALSE
message: FALSE
---

GitHub REPO: <https://github.com/marcelotmiller/ENVS-193DS_spring-2025_final.git>

```{r}
library(tidyverse)
library(flextable)
library(gt)
library(janitor)
library(MuMIn)
library(scales)
library(here)
library(dplyr)
library(readr)
library(lubridate)
library(janitor)
library(DHARMa)
```

\[#1\]

a.)

In part 1, based on wording including correlation and the use of continuous predictor and response variables, we know that a correlation test was used. The most likely test is a Pearson correlation coefficient, which was used to gauge the strength and direction of the proposed linear relationship between the distance from headwater (km) and the annual total nitrogen load (kg yr\^-1).

I part 2, we can assume that an ANOVA test was run to test for significant differences in average nitrogen load (kg yr\^-1) over multiple predictor variables. Other than being a multivariable, we can conclude that an ANOVA test was used because of the specific wording of having no difference in average, and the use of a p-value.

b.)

In addition to mentioning the test type, incorporating a post-hoc comparison, like Turkey's HSD, would specify which predictor variables had significant differences in nitrogen load. Doing so would offer more insight to which predictor variables are worth noting, rather than leaving the ANOVA p-value result of 0.02 left up to the bare interpretation that at least one group differs.

Another detail that can enhance the results is to include an effect size. Including an effect size estimate, such as an eta-squared for the ANOVA test, would help interpreters understand how large the differences that were deemed significant are, and whether the results are to be considered meaningful in actual practice.

c.)

Part 1;

Biological narrative: Nitrogen load trends towards an increase as the distance from the headwater increases, suggesting an ongoing accumulation of nitrogen as it travels downstream.

Statistical summary: A significant positive relationship between the distance from headwater (km) and total annual nitrogen load (kg) was found (test= Pearson correlation, r = (), p= 0.03, alpha= 0.05, df= ().)

Part 2:

Biological narrative: Nitrogen loads are seen to vary across several sources, with specific variables contributing more to the nitrogen accumulation in the San Joaquin River Delta than others. Such findings reveal that there is a disproportionate level of land use contributing to nitrogen pollution in proximity to the watershed.

Statistical summary:

A significant difference in mean annual nitrogen load (kg yr\^-1) between nitrogen source categories (urban land, atmospheric deposition, fertilizer, wastewater treatment, and grasslands) (test = one-way ANOVA, F=(), p= 0.02, alpha= 0.05, df= ()). Post-hoc comparisons reveal that \_ sites are those that differ significantly, corresponding to \_..., effect sizes which show the level of influence.

#2

```{r}
sst_clean <- read.csv("data/sst.csv") |> 
  mutate(data = ymd(date)) |> 
  mutate(year = year(data), 
         month = month(data, label = TRUE, abbr = TRUE)) |> 
  filter(year >= 2018, year <= 2023) |> 
  group_by(site, year, month) |> 
  summarize(mean_monthly_sst = mean(temp, na.rm = TRUE),
            .groups= "drop") |> 
  mutate(month = factor(month, ordered = TRUE))

sst_clean |> 
  slice_sample(n= 5)

str(sst_clean)
```

```{r}
dev.off()

sst_plot_data <- sst_clean |> 
  group_by(year, month) |> 
  summarize(mean_sst = mean(mean_monthly_sst, na.rm = TRUE), 
            .groups = "drop") |> 
  mutate(
    year = factor(year),
    month = factor(month, levels = month.abb, ordered = TRUE)
  )
ggplot(sst_plot_data, aes(x= month, y= mean_sst, group= year, color= factor(year))) + 
  geom_line(size = 0.6) +
  geom_point(size = 2) +
  scale_color_manual(values = c(
    "2018" = "#c6dbef",
    "2019" = "#9ecae1",
    "2020" = "#6baed6",
    "2021" = "#4292c6",
    "2022" = "#2171b5",
    "2023" = "#084594"
  )) +
  labs(
    x = "Month",
    y = "Mean Monthly Sea Surface Temperature (°C)",
    color = "Year"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = c(0.05, 0.90),
    legend.title = element_text(size =11),
    legend.text = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
```

#3\]

a.) In this data set, the 1 represents as a nest box that is occupied by a Swift Parrot, and the 0 represents a box that is not occupied by a Swift Parrot. In other words, the 0 and 1 represent whether or not there was a Swift Parrot in a nest box.

b.) The primary purpose of this study was to analyze the main conservation species of the Swift Parrot, while taking into consideration the competitor species of the Common Starlings and Tree Martins. A nest box occupancy for the Swift Parrot represented a successful recording, and occupancy for other species in nest boxes were recorded as possible negative influences for conservation effectiveness.

c.) The years/seasons that are compared in this study are 2016 and 2019. The season/year of 2016 represents the breeding season including the Swift Parrots, and 2019 represents the season without the swift parrots.

d.)

| Model \# | Season Included (yes/no) | Distance to Forest Edge Included (yes/no) | Model Description |
|---------------|---------------|---------------------|----------------------|
| 1 | No | No | Null model (intercept only) |
| 2 | Yes | Yes | Saturated model (season + distance) |
| 3 | Yes | No | Season only |
| 4 | No | Yes | Distance only |

e.)

```{r}
occdist <- read.csv("data/occdist.csv")
names(occdist) <- make.names(names(occdist))
occdist$season <- as.factor(occdist$season)
model1 <- glm(sp ~ 1, data = occdist, family = binomial)
model2 <- glm(sp ~ season + edge.distance, data = occdist, family = binomial)
model3 <- glm(sp ~ season, data = occdist, family = binomial)
model4 <- glm(sp ~ edge.distance, data = occdist, family = binomial)

```

f.)

```{r}
simulateResiduals(model1) |> plot()
simulateResiduals(model2) |> plot()
simulateResiduals(model3) |> plot()
simulateResiduals(model4) |> plot()

```

g.)

```{r}
model.sel(model1, model2, model3, model4)

```

The best model based on Akaike's Information Criteria is model 2, or the model that includes both season and edge distance as predictor variables in the Swift Parrot nest box occupancy.

h.)

```{r}
newdata <- expand.grid(
  edge.distance = seq(0, 900, by = 50),
  season = c("2016", "2019")
)

predictors <- predict(model2, newdata, type = "link", se.fit = TRUE)
newdata$lfit <- plogis(predictors$fit)
newdata$lower <- plogis(predictors$fit - 1.96 * predictors$se.fit)
newdata$upper <- plogis(predictors$fit + 1.96 * predictors$se.fit)

ggplot(newdata, aes(x = edge.distance, y= lfit, color = season)) +
  geom_line(size = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = season), alpha = 0.2, color = NA) +
  labs(x = "Distance to Forest Edge (m)",
       y = "Predicted Probability of Swift Parrot Occupancy",
       color = "Season", fill = "Season") + 
  theme_minimal(base_size = 13) + 
  theme(panel.grid = element_blank()) +
  scale_color_manual(values = c("2016" = "skyblue", "2019"= "brown2")) + 
  scale_fill_manual(values = c("2016" = "skyblue", "2019"= "brown2"))

```

i.)

Figure 1. Modeled plot predictors showing Swift Parrot next box occupancy as a function of distance from the forest edge in 2019 and 2019. The predicted probabilities from a binomial GLM using season and edge distances as constant predictor variables. Confidence intervals are shown as shaded regions above and below lines. Data source: Stojanovic, Dejan et al. (2021). *Do nest boxes breed the target species or its competitors?*\[Dataset\]. Dryad. <https://doi.org/10.5061/dryad.83bk3j9sb>

j.)

```{r}
testpoints <- expand.grid(
  edge.distance = c(0, 900),
  season = c("2016", "2019")
)

testpreds <- predict(model2, newdata = testpoints, type = "link", se.fit = TRUE)
testpoints$fit <- plogis(testpreds$fit)
testpoints$lower <- plogis(testpreds$fit - 1.96 * testpreds$se.fit)
testpoints$upper <- plogis(testpreds$fit + 1.96 * testpreds$se.fit)

print(testpoints)

```

k.)

We see that the Swift Parrot nest box occupancy was dependent on the season and distance from the forest edge. In the year of 2016, we see that the probability of occupancy was higher near the forest edge as compared to that in 2019, demonstrating the difference in the specie's absence in years.

#4\]
